---
title: "Covid91 vaccine study Final2022"
author: "Madeleine Alseryani"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: "`r Sys.Date()`" 
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
FinalDat <-read.csv(file="final_Data.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- FinalDat  %>% select(-one_of("X"))
FinalData <- FinalData %>%
mutate(across(where(is.character), as.factor))
FinalData$treatment <- relevel(FinalData$treatment,ref="placebo")
#FinalData$infected <- relevel(FinalData$infected,ref="OK")
male = filter(FinalData, sex =="M")
female = filter(FinalData, sex =="F")
gay = filter(FinalData, LGBTQ =="gay")
Druguser = filter(FinalData, DrugUser =="yes")
```

# Introduction

A new  preventative vaccine (covid91) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups by gender( guys and gals), drug use, and sexual orientation. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.


## Males


#### Plot barplots ggplot way

```{r}
dd2 <- male %>% group_by(infected,treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


basicC <- ggplot(dd2,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd2,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```

This bar plot represents the population of males and the relationship of the contraction of covid-19 by either taking a placebo or the vaccine.

The bar plot depicts that about roughly 26% of the male population that got covid-19 took the vaccine whereas, about 68% of the males that got covid-19 took the placebo. For the people who got covid-19, if we compare who took the placebo and who took the vaccine, we can see that the difference is about half. A substantially higher percentage of people who took the placebo contracted covid-19 versus those who got the vaccine itself. This indicates that the
vaccine appears to be pretty effective for the male population that took it. Hence, the vaccine worked for the male population.

We are given two different plots to look at that share the same data. Looking at the first plot, the x-axis represents the individuals that were contacted with covid-19 or those that were not. The y-axis represents the count, which is the number of people that got covid-19 or not. In the second plot, the left two bars of the graph are all the people that got covid-19. The x-axis represents whether a person got covid-19 or not. The y-axis now represents the percentage of people that got covid-19 after taking the the placebo or taking the Vaccine. 


### Numerical results for males

```{r}
table1 <- xtabs(~infected + treatment, data=male)
rowPerc(table1)
colPerc(table1)
```

The numerical results above show the percentages of the male population that took the placebo and contracted covid-19 and those who contracted covid-19 and took the vaccine. About 68% of males that took the placebeo got covid-19 and 32% that took the vaccine got covid-19. The vaccine is clearly effective here in preventing covid-19 among males. Those who took the vaccine were much less likely to get covid-19. The odds of getting covid-19 for this group while taking the vaccine is 2:1.

### Inferential results for males


```{r}
chisq.test(table1)
chisqtestGC(table1)
malefish <- fisher.test(table1)
malefish
```

After running the Fisher's exact test, we always want to look at the given P-value. We acquired a P-value of 2.645e-11. This is a super small number which indicates that there is a significant difference and that these results are not likely or commonly to occur by chance. 

The odds ratio above depicts a ratio of 2.09. Therefore the males that took the vaccine were less likely to contract the virus than those who received the placebo.

```{r}
effect <- malefish$p.value/(malefish$p.value +1)
effectivenessmale = malefish$estimate/(malefish$estimate +1)*100
print(effectivenessmale)
```

```{r}
low=(malefish$conf.int[1]/(malefish$conf.int[1] +1))*100
hi=(malefish$conf.int[2]/(malefish$conf.int[2] +1))*100
options(digits=2)
```

## Effectiveness Confidence Interval `r low`  to `r hi`

Above is the effectiveness for the covid-19 vaccine among the male population. If the effectiveness is above 50% that means that the vaccine is effective. Based on the interval seen above, we can easily conclude that the vaccine is effective in preventing males from contacting covid-19. 

The odds ratio is approximately how effective the vaccine is. In order for the vaccine to be effective, we said that we would need the value to be atleast 50 or higher. The odds ratio for the guys can be seen above as 67.6. The effectiveness confidence interval for the males is 62.51 to 72.51. Based on the interval above, we can conclude that the vaccine is highly effective for the male population.

## Females

#### Plot barplots ggplot way

```{r}
dd2 <- female %>% group_by(infected,treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


basicC <- ggplot(dd2,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd2,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```

This bar plot represents the population of females and the relationship of the contraction of covid-19 by either taking a placebo or the vaccine.

The bar plot depicts that about roughly 40% of the female population that got covid-19 took the vaccine whereas, about 59% of the females that got covid-19 took the placebo. If we compare who took the placebo and who took the vaccine, we can see that the difference is about a little less than half. A larger percentage of people who got covid-19 took the placebo versus those who got the vaccine itself. This indicates that the vaccine appears to be effective for the female population. Even though the effectiveness of the vaccine was not as great as the males effectiveness, it is still effective just not as much. Hence the vaccine for the females works.


### Numerical results for females

```{r}
table1 <- xtabs(~infected + treatment, data=female)
rowPerc(table1)
colPerc(table1)
```

The numerical results above show the percentages of the female population that contracted covid-19 after taking the placebo and the vaccine. About 59% of females got covid-19 after taking the placebo and 41% got covid-19 after taking the vaccine. The vaccine is clearly effective here in preventing covid-19 among females. Those who took the vaccine were less likely to get covid-19. The odds of getting covid-19 while taking the vaccine is 2:1.

### Inferential results for females

```{r}
chisq.test(table1)
chisqtestGC(table1)
femalefish <- fisher.test(table1)
femalefish
```

After running the Fisher's exact test, we always want to look at the given P-value. We acquired a P-value of 9e-04. This is a super small number which indicates that there is a significant difference.

The odds ratio above depicts a ratio of 1.4. Therefore the females that received the vaccine were less likely to contract the virus than those who received the placebo. In other words, the females that contracted covid-19 more were those that took the placebo.

```{r}
effect <- femalefish$p.value/(femalefish$p.value +1)
effectivenessfemale = femalefish$estimate/(femalefish$estimate +1)*100
print(effectivenessfemale)
```

```{r}
low=(femalefish$conf.int[1]/(femalefish$conf.int[1] +1))*100
hi=(femalefish$conf.int[2]/(femalefish$conf.int[2] +1))*100
options(digits=2)
```

## Effectiveness Confidence Interval `r low`  to `r hi`

Above is the effectiveness for the covid-19 vaccine among the female population. If the effectiveness is above 50% that means that the vaccine is effective. Based on the interval seen above, we can easily conclude that the vaccine is effective in preventing females from contacting covid-19. 

The odds ratio is approximately how effective the vaccine is. In order for the vaccine to be effective, we said that we would need the value to be atleast 50 or higher. The odds ratio for the females can be seen above as 59. The effectiveness confidence interval for the females is 53.5 to 64.01. Based on the interval above, we can conclude that the vaccine is effective for the female population.


## LGBTQ

#### Plot barplots ggplot way

```{r}
dd2 <- gay %>% group_by(infected,treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


basicC <- ggplot(dd2,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd2,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```

This bar plot represents the population of LGBTQ individuals and the relationship of the contraction of covid-19 by either taking a placebo or the vaccine.

The bar plot depicts that about roughly 80% of the LGBTQ population that got covid-19 took the vaccine whereas, about 19% of the LGBTQ indivduals that got covid-19 took the placebo. For the LGBTQ indivduals who got covid, if we compare who took the placebo and who took the vaccine, we can see that the difference is very huge. The difference is about 5 or 6X. Therefore, taking the placebo for an LGBTQ person is safer than taking the vaccine. A substantially higher percentage of people who got covid-19 took the vaccine versus those who got the placebo. This indicates that the vaccine is not effective at all. The vaccine is killing this specific group of individuals. Hence, the vaccine did not work for the LGBTQ population.


### Numerical results for LGBTQ

```{r}
table1 <- xtabs(~infected + treatment, data=gay)
rowPerc(table1)
colPerc(table1)
```

The numerical results above show the percentages of the LGBTQ population that contracted covid-19 after taking the placebo and the vaccine. About 13%
of the LGBTQ got covid-19 after taking the placebo and 87% got covid-19 after taking the vaccine. The vaccine is clearly uneffective here in preventing covid-19 among the LGBTQ community. Those who took the vaccine were much more likely to get covid-19. The odds of getting covid-19 while taking the vaccine 0.5:3. This means the chances of contracting covid with the vaccine is enormously high.

### Inferential results for LGBTQ


```{r}
chisq.test(table1)
chisqtestGC(table1)
gayfish <- fisher.test(table1)
gayfish
```

After running the Fisher's exact test, we always want to look at the given P-value. We acquired a P-value of2e-09. This is a super small number which indicates that there is a significant difference and that these results are not likely or commonly to occur by chance. 

The odds ratio above depicts a ratio of 0.15 which is about right according to the data analyzed. Therefore the LGBTQ individuals that got covid-19 received the vaccine and were very likely to contract the virus than those who received the placebo.

```{r}
effect <- gayfish$p.value/(gayfish$p.value +1)
effectivenessgay = gayfish$estimate/(gayfish$estimate +1)*100
print(effectivenessgay)
```


```{r}
low=(gayfish$conf.int[1]/(gayfish$conf.int[1] +1))*100
hi=(gayfish$conf.int[2]/(gayfish$conf.int[2] +1))*100
options(digits=2)
```

## Effectiveness Confidence Interval `r low`  to `r hi`

Above is the effectiveness for the covid-19 vaccine among the LGBTQ population. If the effectiveness is above 50% that means that the vaccine is effective. Based on the interval seen above, we can easily conclude that the vaccine is effective in preventing males from contacting covid-19. 

The odds ratio is approximately how effective the vaccine is. In order for the vaccine to be effective, we said that we would need the value to be atleast 50 or higher. The odds ratio for the guys can be seen above as 13. This is clearly very low and indicates that it will not work. The effectiveness confidence interval for the LGBTQ individuals is 6.65 to 23.78. Based on the interval above, we can conclude that the vaccine is highly effective for the LGBTQ population.


## Druggies

#### Plot barplots ggplot way

```{r}
dd2 <- Druguser %>% group_by(infected,treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


basicC <- ggplot(dd2,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd2,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```

This bar plot represents the population of drug users and the relationship of the contraction of covid-19 by either taking a placebo or the vaccine.

The bar plot depicts that about roughly 85% of the drug user population that got  covid-19 took the vaccine whereas, about 15% of them that got covid-19 took the placebo. For the individuals who got covid-19, if we compare who took the placebo and who took the vaccine, we can see that the difference is very huge. A substantially higher percentage of drug users who took the vaccine contracted covid-19 versus those who got the placebo. This indicates that the
vaccine appears to be killing them and not effective for the drug user population that took it. Hence, the vaccine does not work for the drug user population.


### Numerical results for Drugusers

```{r}
table1 <- xtabs(~infected + treatment, data=Druguser)
rowPerc(table1)
colPerc(table1)
```

The numerical results above show the percentages of the drug user population that contracted covid-19 after taking the placebo and the vaccine. About 15% 
of drug users got covid-19 after taking the placebo and 
85% got covid-19 after taking the vaccine. The vaccine is clearly noneffective here in preventing covid-19 among drug users. Those who got covid-19 took the vaccine. They were much more likely to get covid-19. The odds of getting covid-19 while taking the vaccine is 0.6:3.3. Therefore, for this specific group of people, taking the placebo is more effective than taking the vaccine.

### Inferential results for Drugusers


```{r}
chisq.test(table1)
chisqtestGC(table1)
Druguserfish <- fisher.test(table1)
Druguserfish
```

After running the Fisher's exact test, we always want to look at the given P-value. We acquired a P-value of 3e-08. This is a super small number which indicates that there is a significant difference and that these results are not likely or commonly to occur by chance. 

The odds ratio above depicts a ratio of 0.18 which is about right according to the data analyzed. Therefore the drug users that received the vaccine were very much more likely to contract the virus than those who received the placebo.

```{r}
effect <- Druguserfish$p.value/(Druguserfish$p.value +1)
effectivenessDruguser = Druguserfish$estimate/(Druguserfish$estimate +1)*100
print(effectivenessDruguser)
```


```{r}
low=(Druguserfish$conf.int[1]/(Druguserfish$conf.int[1] 
+1))*100
hi=(Druguserfish$conf.int[2]/(Druguserfish$conf.int[2]
+1))*100
options(digits=2)
```

## Effectiveness Confidence Interval `r low`  to `r hi`

Above is the effectiveness for the covid-19 vaccine among the drug user population. If the effectiveness is above 50% that means that the vaccine is effective. Based on the interval seen above, we can easily conclude that the vaccine is effective in preventing males from contacting covid-19.

The odds ratio is approximately how effective the vaccine is. In order for the vaccine to be effective, we said that we would need the value to be atleast 50 or higher. The odds ratio for the drug users can be seen above as 15. This is clearly very low and indicates that it will not work. The effectiveness confidence interval for the drug user individuals is 7.11 to 26.78. Based on the interval above, we can conclude that the vaccine is not effective for the drug user population.


## Conclusion

In conclusion, the male population was the group that had the best effectiveness of the vaccine, followed by the females, followed by the drug users, and lastly the worst group being the LGBTQ population. In this experiment, I was able to look at bar plots and analyze them , look at numerical results, inferential results, run a fisher's exact test and analyze the P-value, and then run the confidence interval to see the effectiveness of the vaccine. After analyzing this data for each specific group given, I can conclude that the male population came in first place for having the most effective vaccine in preventing covid-19 and  the LGBTQ population had the worst effectiveness of the vaccine. 
